Equalizer\\Programming and User Guide

### 2011-11-18 Developer/ProgrammingGuide/paper.tex
Add auto-config section to User Guide

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications using
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.8.3 for Equalizer 1.1.3+git \\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.8.1   & Mar 25, 2011 & Minor Revision\\
    1.8.2   & Apr 04, 2011 & Added \sref{sFocusDistance}\\
    1.8.3   & Aug 10, 2011 & Rewrote \sref{sHello}\\
    \hline
  \end{tabularx}}
\clearpage

@@

\section{\label{sConfig}Writing Configurations}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewLB.pdf}
  {\caption{\label{fViewLoadBalancing}\small Cross-Segment
      Load-Balancing for two Segments using eight GPU's}}
\end{wrapfigure}
Depending on the model position and data structure, each segment of a
multi-display system has a different rendering load. The segment with
the biggest load determines the overall performance when using a static
assignment of resources to segments. The view equalizer analyzes the
load of all segments, and adjusts the resource usage each frame. It
equalizes the load on all segments of a view.

@@

\begin{wrapfigure}{r}{.382\textwidth} 
    \includegraphics[width=.382\textwidth]{images/viewLB.png}
    {\caption{\label{fViewLoadBalancing2}\small Cross-Segment
        Load-Balancing for a CAVE}}
\end{wrapfigure}
The view equalizer can also use resources from another display resource,
if this resource has little rendering load by itself. It is therefore
possible to improve the rendering performance of a multi-display system
without any additional resources. This is particularly useful for
installations with a higher number of displays where the rendering load
is typically in a few segments only, e.g., for a CAVE.

\fig{fViewLoadBalancing2} shows cross-usage for a five-sided CAVE driven
by five GPU's. The front and left segments show the model and have a
significant rendering load. The view equalizer assigns the GPU's from
the top, bottom and right wall for rendering the left and front wall in
this configuration.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/dfr.pdf}
  {\caption{\label{fDFR}Dynamic Frame Resolution}}
\end{wrapfigure}
Dynamic Frame Resolution (DFR) trades rendering performance for visual
quality. The rendering for a channel is done at a different resolution
than the native channel resolution to keep the framerate constant. The
DFR equalizer adjusts the zoom of a channel, based on the target and
current framerate. It is typically used for fill-rate bound
applications, such as volume rendering and ray-tracing.

@@

\subsection{Debugging}
If your configuration does not work, simplify it as much as possible
first. Normally this means that there is one server, application and render
client. Failure to launch a cluster configuration often is caused by one of the
following reasons:
\begin{itemize}
\item A firewall is blocking network connections.
\item The render client can't access the GPU's on the remote host. Set up your
  X-Server or application rights correctly. Log into the remote machine using
  the launch command and try to execute a simple application, e.g.,
  \textsf{glxinfo -display :0.1}.
\item The server does not find the prelaunched render client. Verify that the
  client is listening on the correct IP and port, and that this IP and port are
  reachable from the server host.
\item The server cannot launch a render client. Check the server log for the
  launch command used, and try to execute a simple application from the server
  host using this launch command. It should run without user interaction. Check
  that the render client is installed in the correct path. Pay attention to the
  launch command quotes used to separate arguments on Windows. Check that the
  same software version, including Equalizer, are installed on all machines.
\item A client can't connect back to the application. Check the client log, this
  is typically caused by a misconfigured host name resolution.
\end{itemize}

>>>>

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications using
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.8.4 for Equalizer 1.1.5+git \\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.8.1   & Mar 25, 2011 & Minor Revision\\
    1.8.2   & Apr 04, 2011 & Added \sref{sFocusDistance}\\
    1.8.3   & Aug 10, 2011 & Rewrote \sref{sHello}\\
    1.8.4   & Nov 18, 2011 & Added \sref{sAutoConfig}\\
    \hline
  \end{tabularx}}
\clearpage

@@

\section{\label{sConfig}Configuring Equalizer Clusters}

\subsection{\label{sAutoConfig}Auto-Configuration}

\subsubsection{GPU Service Discovery}
Equalizer uses the GPU-SD library for discovering local and remote GPUs. Based
on this information, a typical configuration is dynamically compiled at
application start. The auto-configuration uses the same format as the static
configuration files, and can therefore be used to create template configurations
by running a test application and copying the generated configuration.

Please note that the ZeroConf module is optional and requires a ZeroConf
implementation when Equalizer is compiled.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/gpu-sd.pdf}
  {\caption{\label{fGPUSD}GPU discovery for auto-configuration}}
\end{wrapfigure}
GPU-SD contains three components: a core library, local and remote discovery
modules as well as a ZeroConf daemon. The daemon uses the appropriate module to
query local GPUs and announces them using ZeroConf on the local network.

Equalizer uses the library with the dns\_sd discovery module to gather the
information announced by all daemons on the local network, as well as the local
discovery modules for standalone auto-configuration. \fig{fGPUSD} shows how
GPU-SD is used to discover four remote GPUs on two different machines to use
them to render a single view on a laptop.

\subsubsection{Usage}

On each node contributing to the configuration, install and start the GPU-SD
daemon. If multiple, disjoint configurations are used on the same network,
provide the session name as a parameter when starting the daemon. Verify that
all GPUs are visible on the application host using the \textsf{gpu\_sd\_list}
tool. When starting the application, use the command-line parameter
\textsf{--eq-config}:

\begin{itemize}
\item The default value is \textsf{local} which uses all local GPUs queried
  using the cgl, glx or wgl GPU-SD module.
\item \textsf{--eq-config sessionname} uses the dns\_sd ZeroConf module of
  GPU-SD to query all GPUs in the subnet for the given session. The default
  session of the GPU-SD daemon is \textsf{default}. The ZeroConf hostname is
  used to connect the nodes.
\item \textsf{--eq-config filename.eqc} loads a static configuration from the
  given ascii file. The following sections describe how to write configuration
  files.
\end{itemize}

The auto-configuration creates one display window on the local machine, and one
off-screen channel for each GPU. The display window has one full-window channel
used as an output channel for a single segment. It combines all GPUs into a
scalability config with different layouts for each of the following scalability
modes:

\begin{description}
\item [2D] A dynamically load-balanced sort-first configuration using all GPUs
  as sources.
\item [Simple] A no-scalability configuration using only the display GPU for
  rendering.
\item [Static DB] A static sort-last configuration distributing the range evenly
  across all GPUs.
\item [Dynamic DB] A dynamically load-balanced sort-last configuration using all
  GPUs as sources.
\end{description}

The ZeroConf hostname is used to configure the node, that is, the launch command
has to be able to resolve this hostname for starting the render client
processes. Given a properly set up ZeroConf installation, the default launch
command \textsf{ssh} should be able to start the render client correctly.

@@

Depending on the model position and data structure, each segment of a
multi-display system has a different rendering load. The segment with
the biggest load determines the overall performance when using a static
assignment of resources to segments. The view equalizer analyzes the
load of all segments, and adjusts the resource usage each frame. It
equalizes the load on all segments of a view.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewLB.pdf}
  {\caption{\label{fViewLoadBalancing}\small Cross-Segment
      Load-Balancing for two Segments using eight GPU's}}
\end{wrapfigure}
The view equalizer can also use resources from another display resource,
if this resource has little rendering load by itself. It is therefore
possible to improve the rendering performance of a multi-display system
without any additional resources. This is particularly useful for
installations with a higher number of displays where the rendering load
is typically in a few segments only, e.g., for a CAVE.

\begin{wrapfigure}{l}{.382\textwidth} 
    \includegraphics[width=.382\textwidth]{images/viewLB.png}
    {\caption{\label{fViewLoadBalancing2}\small Cross-Segment
        Load-Balancing for a CAVE}}
\end{wrapfigure}
\fig{fViewLoadBalancing2} shows cross-usage for a five-sided CAVE driven
by five GPU's. The front and left segments show the model and have a
significant rendering load. The view equalizer assigns the GPU's from
the top, bottom and right wall for rendering the left and front wall in
this configuration.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-3ex}\includegraphics[width=.618\textwidth]{images/dfr.pdf}
  {\caption{\label{fDFR}Dynamic Frame Resolution}}
\end{wrapfigure}
Dynamic Frame Resolution (DFR) trades rendering performance for visual
quality. The rendering for a channel is done at a different resolution
than the native channel resolution to keep the framerate constant. The
DFR equalizer adjusts the zoom of a channel, based on the target and
current framerate. It is typically used for fill-rate bound
applications, such as volume rendering and ray-tracing.

@@

\subsection{Debugging}
If your configuration does not work, simplify it as much as possible
first. Normally this means that there is one server, application and render
client. Failure to launch a cluster configuration often is caused by one of the
following reasons:
\begin{itemize}
\item A firewall is blocking network connections.
\item The render client can't access the GPU's on the remote host. Set up your
  X-Server or application rights correctly. Log into the remote machine using
  the launch command and try to execute a simple application, e.g.,
  \textsf{glxinfo -display :0.1}.
\item The server does not find the prelaunched render client. Verify that the
  client is listening on the correct IP and port, and that this IP and port are
  reachable from the server host.
\item The server cannot launch a render client. Check the server log for the
  launch command used, and try to execute a simple application from the server
  host using this launch command. It should run without user interaction. Check
  that the render client is installed in the correct path. Pay attention to the
  launch command quotes used to separate arguments on Windows. Check that the
  same software versions, including Equalizer, are installed on all machines.
\item A client can't connect back to the application. Check the client log, this
  is typically caused by a misconfigured host name resolution.
\end{itemize}
